---
title: 'Extra Credit: More JSON Practice'
author: "Kristin Lussi"
date: "2023-11-05"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this extra credit, we will load the Nobel Prize data using an API from nobelprize.org. We will then ask 4 interesting questions and provide answers to these questions using the data.

## Load Packages

```{r, warning = FALSE, message = FALSE}
library(jsonlite)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(httr)
library(gt)
```


## Load and Clean Data

```{r, warning = FALSE, message = FALSE}
prize_url <- "http://api.nobelprize.org/v1/prize.json"
laureate_url <- "http://api.nobelprize.org/v1/laureate.json"

nobelPrize <- fromJSON(prize_url)

nobelLaureate <- fromJSON(laureate_url)

nobelPrize <- nobelPrize$prizes %>%
  unnest_wider(laureates) %>%
  unnest(id, firstname, surname, motivation, share)

nobelPrize_df <- bind_rows(nobelPrize)
nobelLaureate_df <- bind_rows(nobelLaureate)

# join the prize and laureate data frames by id
final_df <- inner_join(nobelPrize_df, nobelLaureate_df, by = "id") %>%
  select(-c("firstname.y", "surname.y", "prizes", "overallMotivation")) %>%
  as.data.frame()

head(final_df)
```

## Questions

### What is the most common country of birth for Nobel Prize winners in the past 5 years?

```{r}
final_df %>%
  group_by(bornCountry) %>%
  filter(year %in% 2018:2023) %>%
  filter(n() > 1) %>% 
  filter(!is.na(bornCountry)) %>%
  ggplot() +
  geom_bar(aes(x = bornCountry, fill = bornCountry)) +
  ggtitle("Birth Countries of Nobel Prize Winners",
          "From 2018 to 2023") +
  ylab("Count") +
  xlab("Birth Country") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1))
  
```

From the above graph, we can see that the most common country of birth for Nobel prize winners in the past 5 years is the United States. I thought this was interesting, so I did some outside research. One theory as to why the majority of Nobel prizes go to Americans is due funding and academic freedom. An article on InsideScience.org says, "Since the mid-20th century, the United States has spent a tremendous amount on fundamental or "basic" research, not forcing scientists to work on projects with an immediate application as the goal."

### As we know, a Nobel Prize can be split between up to three people. What is the most commonly split prize category (in the past 5 years)?

```{r}
final_df %>%
  group_by(year, category) %>%
  filter(year %in% 2018:2023) %>%
  ggplot() +
  geom_bar(aes(x = category, fill = year), position = "dodge") +
  ggtitle("Nobel Prize Categories",
          "From 2018 to 2023") +
  ylab("Count") +
  xlab("Prize Category") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Year")
```

From the above graph, we can see that the physics prize has been split between three winners every year in the past 5 years. Chemistry and economics are the next most commonly split prize categories. It is interesting to see that the literature prize has not been split at all in the past 5 years. 

### How many Nobel Prize winners were born in the 1990's?

```{r}
final_df %>%
  mutate(yearBorn = substr(born, 1, 4)) %>%
  filter(yearBorn %in% 1990:1999) %>%
  # filter out organizations
  filter(!is.na(surname.x)) %>%
  as.data.frame()
```

From the above data frame, we can see that there have been 2 Nobel prize winners who were born in the 1990's. Both won the Nobel peace prize and both are female. 

### What is the gender split of Nobel Prize winners over the past century?
```{r}
final_df %>%
  mutate(decade = paste0(substr(year, 1, 3), "0s")) %>%
  group_by(decade, gender) %>%
  ggplot() +
  geom_bar(aes(x = decade, fill = gender), position = "dodge") +
  ggtitle("Nobel Prizes by Gender and Decade") +
  xlab("Decade") +
  ylab("Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Gender")
```

```{r}
final_df %>%
  mutate(decade = paste0(substr(year, 1, 3), "0s")) %>%
  group_by(decade) %>%
  mutate(percent_male = sum(gender == "male") / n(),
         percent_female = sum(gender == "female") / n(),
         total = n()) %>%
  select(decade, percent_male, percent_female) %>%
  distinct(decade, .keep_all = TRUE) %>%
  mutate_at(vars(matches("percent_")), ~round(., 2)) %>%
  gt() %>%
  tab_header(title = "Nobel Prizes by Gender") %>%
  cols_label(percent_male = "Percent Male", percent_female = "Percent Female")
  
```


From the above graph, we can see that men have won the majority of Nobel prizes every decade since the commencement of the Nobel Prize. However, from the table we can see that the percentage of female prize winners has been increasing each decade since the 1990's. 

## Sources

https://nobelprize.readme.io/reference/laureate

https://www.insidescience.org/news/why-do-so-many-americans-win-nobel-prize#:~:text=According%20to%20experts%2C%20it%27s%20strong,and%20patience%20to%20see%20results.
