---
title: "DATA 607 Assignment 9"
author: "Kristin Lussi"
date: "2023-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
api_key <- "ViibaPGCqmcdLaPRIhGVQNpcZCc0Td3x"
```

## Introduction

In this assignment, we will load the New York Times Best Sellers list using an API key. Then, we will use the data to answer the following question: *Which book category is ranked highest on the NYT Best Sellers list on average?*

## Load Packages

First, we load the necessary packages:

```{r, warning = FALSE, message = FALSE}
library(jsonlite)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Retrieve JSON from API

Next, we will retrieve the JSON data from the API url, using my API key.

```{r}
# specify url 
api_url <- paste0("https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?api-key=", api_key)

# use the fromJSON function from the jsonlite package to parse the JSON data
bookReviews <- fromJSON(api_url)
```

## Unnest Data

Next, we will unnest the data to create a clean data frame.

```{r, warning = FALSE, message = FALSE}
bookReviews_df <- as.data.frame(bookReviews$results)

bookRankings <- bookReviews_df %>%
  unnest_wider(isbns) %>%
  unnest_wider(ranks_history) %>%
  unnest_wider(reviews) %>%
  unnest(display_name) %>%
  unnest(rank) %>%
  # select only title, author, display name, and book ranking
  select(title, author, rank, display_name) %>%
  # select only one distinct title, no duplicates
  distinct(title, .keep_all = TRUE)%>%
  as.data.frame()

bookRankings
```

## Analysis

Now, we will answer the question: *Which book category is ranked highest on the NYT Best Sellers list on average?*

We will first calculate the average ranking for each category:
```{r}
# calculate average ranking for each category
avgRankings <- bookRankings %>%
  group_by(display_name) %>%
  summarise(AverageRanking = mean(rank)) %>%
  as.data.frame()

colnames(avgRankings) <- c("Category", "AverageRanking")

avgRankings
```

Now, we will visualize the results using a bar plot:

```{r}
# plot average rankings for each category
ggplot(avgRankings) +
  geom_bar(aes(x = Category, y = AverageRanking, fill = Category), stat = "identity") +
  coord_flip() +
  ggtitle("Average Rankings for Each Book Category",
          "From New York Times Best Sellers List") +
  ylab("Average Ranking") +
  xlab("Book Category") +
  theme(axis.text.y = element_text(angle = 20, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")
```

## Conclusion

As we know, the lower number a book is ranked, the higher it is on the New York Times Best Sellers list. So, the category with the lowest average ranking is more often ranked higher on the Best Sellers list. According to our graph, this category is "Combined Print & E-Book Nonfiction". 

## Sources

"New York Times: Best Sellers"
The New York Times, https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json
